Mental simulation as Bayesian Quadrature

### 2013-05-31 man/nips-2013.tex
Add Tom intro edits

Consider the images on the left in Figure
\ref{fig:mental-rotation}. In each panel, are the two depicted objects
identical (except for a rotation), or distinct? When presented with
this mental rotation task, people default to a strategy in which they
visualize one object rotating until it is congruent with the other
\cite{Shepard1971}. There is strong evidence for such \textit{mental
  imagery} or \textit{mental simulation}: we can imagine
three-dimensional objects in our minds and manipulate them, to a
certain extent, as if they were real \cite{Kosslyn:2009tj}.  However,
the use of mental simulation is predicated on determining appropriate
parameters to give the simulation, and people's cognitive constraints
may furthermore place limitations on the duration or precision of
simulation. One hypothesis for how these issues are handled argues
that people use a rational solution, meaning that it is optimal under
existing constraints \cite{Lieder:2012wg,Vul:2009wy,Griffiths2012a},
such as processing speed, memory, or perceptual acuity.

In the case of the classic mental rotation task, we might ask: in what
direction should the object be rotated?  What are the requirements for
``congruence''? When should one stop rotating and accept the
hypothesis that the objects are different? We investigate several
computational solutions to this task and qualitatively compare them to
general mental rotation results. Of particular interest is a model
which actively reasons about the amount of evidence it has, and
performs mental rotations until it reaches a certain level of
confidence. We contrast this model with simpler approaches which are
less computationally demanding, but which employ less flexible
strategies.

>>>>

One of the challenges of solving any computational problem is
determining how best to use the available computing resources. For
example, a computer can render complex graphics faster by recognizing
that this kind of computation should be carried out by a specialized
graphics processor. The same challenge arises in designing an
intelligent agent: how should the agent make best use of its computing
resources? Recent research on rational models of human cognition has
provided insight into the nature of the computational problems that
human beings need to solve (e.g., \cite{something}), but leaves open
the question of how people allocate their resources in solving those
problems. In this paper, we take a step towards addressing this
question, applying rational analysis (in the spirit of
\cite{marr,anderson, shepard}) to one aspect of human metacognition:
the use of mental simulation.

Consider the images on the left in Figure
\ref{fig:mental-rotation}. In each panel, are the two depicted objects
identical (except for a rotation), or distinct? When presented with
this mental rotation task, people default to a strategy in which they
visualize one object rotating until it is congruent with the other
\cite{Shepard1971}. There is strong evidence for such mental
simulation: we can imagine three-dimensional objects in our minds and
manipulate them, to a certain extent, as if they were real
\cite{Kosslyn:2009tj}.  However, the use of mental simulation is
predicated on determining appropriate parameters to give the
simulation, analogous to determining exactly what computation should
be passed to a graphics processor.  In the case of the classic mental
rotation task, we might ask: How do people know which way to rotate
the object?  When should one stop rotating and accept the hypothesis
that the objects are different?

Recent work in cognitive science has shown how problems of allocating
cognitive resources to solving computational problems can be analyzed
using the methods of statistical decision theory
\cite{Lieder:2012wg,Vul:2009wy}. We apply this ``rational
metacognition'' approach to the problem of mental rotation. We
investigate several computational solutions to this task and
qualitatively compare them to human mental rotation performance. In
particular, we argue that performing mental rotation can be framed as
an integration problem, with the direction of rotation becoming a
statistical design problem (or in machine learning parlance, a problem
of active learning). We show that recent work on methods for Bayesian
quadrature \cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}
provides a way to solve this problem, outperforming several simpler
heuristics for determining both the direction and extent of rotation.
